Distribution-oblivious Programming with Dish

### 2019-10-02 pldi/draft.tex
Update intro

To understand this sharp contrast, consider calculating frequencies of terms within a set of input files.
For a developer with \emph{one} computer, this is achievable with a short \unix
pipeline~\cite{bentley1986literate}:

% \begin{lstlisting}[language=sh,float=h,numbers=none]
% cat doc.ms |                   # print input file
% groff -t -e -mandoc -Tascii |  # remove formatting
% col -bx |                      # remove backspaces 
% tr A-Z a-z |                   # convert to lower case
% tr -d '[:punct:]' |            # remove punctuation
% sort |                         # put words in alphabetical order
% unique |                       # remove duplicate words
% comm -13 /usr/share/dict -     # report words not in dictionary 
% \end{lstlisting}
% greece; stops
% AWS policy specification
\begin{lstlisting}[language=sh, float=h, numbers=none, escapeinside={($}{$)}]
cat * | tr -cs A-Za-z`\n` | tr A-Z a-z | (($$p_1$$))
  sort | uniq -c | sort -rn | head 5 > out
\end{lstlisting}

Combining several features, Unix makes small tasks easy to express;
  composition allows the composition of general primitives
Key elements of \unix are the ability to stitch together programs written in different languages, the abstraction of a file system as a set of resident streams, as well as naming---the ability to resolve command, file names within a global context.
Under the hood, the \unix kernel buffers results, concurrency control

@@

Unfortunately, 
A developer with many computers has to choose between three options, all of which require significant manual effort.
The most popular is to leverage domain-specific ; this requires rewriting portions of the program in a new language and use operations

% If developers have already expressed their computation as a Unix pipeline, they should not have to manually rewrite the program in other environment---\eg Hadoop, Spark---to exploit distribution.
For simple, one-off pipelines that take a few minutes to compose this cost---order of magnitude higher than---becomes unjustifiable.
For many legacy pipelines  
Could the generation and execution of a distributing version of the pipeline be automated?

The key insight behind our work is that Unix pipelines already provide most of the information for distributing a computation.
Indeed, the primary reason behind $p_1$'s succinctness is that the pipeline is a domain-specific language for describing operations on streams.
Primitives 
\sys enables \emph{distribution-oblivious programming}: 
  multi-order of magnitude speedups with correctness guarantees and without any developer effort.

>>>>

To understand this contrast, consider calculating term frequencies over a set of input files.
For a user with \emph{one} computer, this is achievable with a short \unix pipeline~\cite{bentley1986literate}:

% \begin{lstlisting}[language=sh,float=h,numbers=none]
% cat doc.ms |                   # print input file
% groff -t -e -mandoc -Tascii |  # remove formatting
% col -bx |                      # remove backspaces 
% tr A-Z a-z |                   # convert to lower case
% tr -d '[:punct:]' |            # remove punctuation
% sort |                         # put words in alphabetical order
% unique |                       # remove duplicate words
% comm -13 /usr/share/dict -     # report words not in dictionary 
% \end{lstlisting}

\begin{lstlisting}[language=sh, float=h, numbers=none, escapeinside={($}{$)}]
cat * | tr -cs A-Za-z\n | tr A-Z a-z |    ($$[p_1]$$)
  sort | uniq -c | sort -rn | head 5 > out
\end{lstlisting}

Combining several features, Unix makes small tasks easy to express:
  $p_1$ creates a character stream, breaks it down to words, transliterates them lower-case, sorts them to group duplicates, counts and replaces duplicates, sorts them numerically and in reverse order, picks the top five results and writes them to a file \ttt{out}.
% composition allows the composition of general primitives
Key elements of \unix are the ability to compose programs written in different languages, the abstraction of a file system as a set of resident streams, a small but extensible library of commands, and the ability to resolve names within a global context.
Under the hood, the \unix kernel buffers results, synchronizes processing stages, and generally orchestrates the computation.

@@

%      dish      ignis, mozart    hadoop, spark, ...      erlang, elixir
%       \/           \   /             \     /                \   /
% auto <-------------------------------------------------------------> manual
%                      [rewrite (not write) cost]

For a user with many computers, however, rewriting this pipeline to exploit distribution requires a significant cost in terms of manual effort.
% The scope of such rewrites, and therefore the cost of manual effort, can vary considerably.
Such cost is at best moderate when pipelines can be expressed in domain-specific frameworks---for example, Hadoop and Spark provide a few purely functional primitives that 
As long as the program fits into the model this is only moderately difficult;
For simple, one-off pipelines that take a few minutes to compose this cost---order of magnitude higher than---becomes unjustifiable.

More realistic programs, such as the pipelines presented in this paper, require distributed programming language or a distributed operating system---\eg Plan9's \ttt{rc} shell.
For many legacy pipelines  
Could the generation and execution of $p_1$'s distributing version be fully and correctly automated?

 has to choose between three options, all of which require significant manual effort.
The most invasive is to rewrite the entire program in a distributed programming language
The most popular is to leverage

Use a distributed programming language, even higher rewrites
Or use some form of annotations

 this requires rewriting portions of the program in a new language and use operations

In the simplest case, annotations

Often, they only focus on a few parts of the system---for example, upgrading to a distributed storage layer.
More rarely, companies rewrite entire systems (\eg Twitter's Ruby-to-Scala rewrite~\cite{twitter}), a process that is notoriously difficult under schedule constraints and competitive pressures~\cite{rewrite1, rewrite2}.
The manual effort is expensive, and can introduce new bugs, cascading changes, or regressions of previously fixed performance issues, especially since software today makes extensive use of third-party modules~\cite{breakapp:plos:2017}.

% If developers have already expressed their computation as a Unix pipeline, they should not have to manually rewrite the program in other environment---\eg Hadoop, Spark---to exploit distribution.

The key insight behind our work is that \unix pipelines already provide most of the information for distributing a computation.
Indeed, the primary reason behind $p_1$'s succinctness is that the pipeline is a domain-specific language for describing operations over streams.
Primitives 
\sys enables \emph{distribution-oblivious programming}: 
  multi-order of magnitude speedups with correctness guarantees and without any developer effort.
